### 2019-11-26 paper/poster/poster.tex
Add first version of poster

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% baposter Landscape Poster
% LaTeX Template
% Version 1.0 (15/5/13)
%
% Created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a0paper,portrait,fontscale=0.3, margin=2cm]{baposter}

\usepackage[utf8]{inputenc}
\usepackage[font=normalsize,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\captionsetup{justification=centering}
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{lmodern}
\usepackage{relsize} % Used for making text smaller in some places

% Other packages
\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}}
\def\imagetop#1{\vtop{\null\hbox{#1}}}
\usepackage{adjustbox}
\usepackage{subfig}
\usepackage{floatrow} % sidecapfloat
\usepackage{setspace}
\usepackage{xspace}
\usepackage{noReferences}
\usepackage{caption}
\usepackage{tikz}
\usepackage[customcolors]{hf-tikz}
%\usepackage{subcaption}
\usepackage{epsf}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{float}
\usepackage{cases}
\usepackage{verbatim}
\usepackage{array}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{multirow}
\usepackage{listings}
%\usepackage[authoryear]{natbib}
\usepackage{authblk}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{eqparbox}
\usepackage{xcolor}
\usepackage[backend=biber, citestyle=authoryear, maxcitenames=2, maxbibnames=2]{biblatex}
\usepackage{lipsum}
\usepackage{svg}
\usepackage{cleveref}
\usepackage{tikz}
\usepackage{xspace}
\usetikzlibrary{arrows,automata}
\usepackage{array, makecell}
\usepackage{pgfplots}
\usetikzlibrary{intersections}

\bibliography{../budgeted_rl.bib}

\input{style}
\input{../mathdef}

\begin{document}

%\background{ % Set the background to an image (background.pdf)
%\begin{tikzpicture}[remember picture,overlay]
%\draw (current page.north west)+(-2em,2em) node[anchor=north west]
%{\includegraphics[height=1.1\textheight]{background.png}};
%\end{tikzpicture}
%}
\begin{poster}{
grid=false,
borderColor=bordercol, % Border color of content boxes
headerColorOne=headercol1, % Background color for the header in the content boxes (left side)
headerColorTwo=headercol2, % Background color for the header in the content boxes (right side)
headerFontColor=headerfontcol, % Text color for the header text in the content boxes
boxColorOne=boxcolor, % Background color for the content in the content boxes
headershape=roundedright, % Specify the rounded corner in the content box headers
%headerfont=\Large\sf\bf, % Font modifiers for the text in the content box headers
headerfont=\Large\bf\textsc, %Sans Serif
textborder=rectangle,
background=none,
headerborder=open, % Change to closed for a line under the content box headers
boxshade=plain,
textfont={\setlength{\parindent}{0.0em}\sffamily},
headerheight={0.05\textheight},
eyecatcher=true
%columns=5
}
%
%----------------------------------------------------------------------------------------
%	Title and authors
%----------------------------------------------------------------------------------------
%
{
\includegraphics[width=15em]{./img/companies}
}
{
Social Attention for Autonomous Decision-Making in Dense Traffic
}
{
Edouard Leurent, Jean Mercat
\vspace{-2\baselineskip}
}
{
\includegraphics[width=12em]{./img/inria_sc}
}

\setlength{\colheight}{0.92\textheight}

%----------------------------------------------------------------------------------------
%	Motivation
%----------------------------------------------------------------------------------------

\headerbox{\textsc{Motivation}}{name=motivation,span=1,column=0,row=0}{

\textbf{Behavioural Planning}

\textbf{Reinforcement Learning}

\begin{align*}
V^\pi(s) &\eqdef \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(s_t, a_t)\right]\\
Q^\pi(s, a) &\eqdef R(s, a) + \gamma \expectedvalueover{s'\sim P(s'|s, a)} V^\pi(s')
\end{align*}

The optimal action-value function $Q^* =  \max_\pi Q^\pi(s)$ satisfies the Bellman Optimality Equation:
\begin{equation*}
Q^*(s, a) = \expectedvalueover{s'\sim P(s'|s, a)} \max_{a'\in A} \left[R(s, a) + \gamma Q^*(s', a')\right]
\end{equation*}

}

%----------------------------------------------------------------------------------------
%	Setting
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Setting}}{name=setting,span=1,column=0,row=0,below=motivation}{

\textbf{State representations for social interactions}

\begin{align*}
s = \left( s_i \right)_{i \in [0, N]}
\end{align*}
where
\begin{align*}
s_i = \begin{bmatrix}
x_i & y_i & v^x_i & v^y_i & \cos\psi_i & \sin \psi_i
\end{bmatrix}^T
\label{eq:coordinates}
\end{align*}

\begin{center}
	\includegraphics[width=0.35\textwidth]{../img/coordinates}

	List of features representation
\end{center}

\textbf{Requirements for function approximation}

\begin{enumerate}
	\item Permutation Invariance

\begin{align*}
\forall\tau \in \mathfrak{S}_N, \qquad &\pi(\cdot|(s_0, s_1,\dotsc,s_N)) = \\ &\pi(\cdot|(s_0, s_{\tau(1)},\dotsc,s_{\tau(N)}))
\end{align*}

\item Variable size

\end{enumerate}

\begin{center}
	\includegraphics[width=0.35\textwidth]{../img/map}

	Occupancy grid representation
\end{center}

}

%----------------------------------------------------------------------------------------
%	Model Architecture
%----------------------------------------------------------------------------------------
\headerbox{Model Architecture}{name=archi,column=1,span=2}{

\textbf{Self-Attention}
\begin{equation}
\text{output}=\underbrace{\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)}_{\text{attention matrix}}V
\label{eq:selfattention}
\end{equation}

\begin{center}

\begin{tikzpicture}
	\node(X1){ego};
	\node[below of=X1, node distance=0.7cm](X2){vehicle$_{1}$};
	\node[below of=X2, node distance=0.6cm](X3){$\vdots$};
	\node[below of=X3, node distance=0.7cm](X4){vehicle$_{N}$};

	\node[draw, right of=X1, node distance=1.6cm, rectangle](ENC1){Encoder};
	\node[draw, right of=X2, node distance=1.6cm, rectangle](ENC2){Encoder};
	\node[below of=ENC2, node distance=0.6cm](ENC3){$\vdots$};
	\node[draw, right of=X4, node distance=1.6cm, rectangle](ENC4){Encoder};

	\path (X1) edge (ENC1);
	\path (X2) edge (ENC2);
	\path (X4) edge (ENC4);

	\node[draw, rectangle, right of=ENC1, node distance=1.6cm, below=-0.4cm, fill=white](TRANS3){\rotatebox{90}{ Ego-attention }};
	\node[draw, rectangle, right of=ENC1, node distance=1.5cm, below=-0.3cm, fill=white](TRANS2){\rotatebox{90}{ Ego-attention }};
	\node[draw, rectangle, right of=ENC1, node distance=1.4cm, below=-0.2cm, fill=white](TRANS1){\rotatebox{90}{ Ego-attention }};

	\draw (ENC1.east) -| (TRANS1.west);
	\draw (ENC2.east) -| (TRANS1.west);
	\draw (ENC4.east) -| (TRANS1.west);

	\node[draw, right of=ENC1, node distance=2.8cm, rectangle](DEC1){Decoder};

	\draw (TRANS1.east) |- (DEC1.west);

	\node[right of=DEC1, node distance=1.2cm](Y1){Q};

	\draw (DEC1.east) -- (Y1.west);
\end{tikzpicture}
\begin{tikzpicture}[scale=0.7, every node/.style={scale=0.7}]
\node(X1){ego encoding};
\node[below of=X1, node distance=2cm](X2){$\vdots$};
\node[below of=X2, node distance=2cm](X3){vehicle$_{N}$ encoding};

\coordinate[right of= X1, node distance=2cm](X1b){};

\draw (X1) -- (X1b);

\node[draw, right of=X1b, node distance=1cm](LK1){$L_{k}$};
\node[draw, below of=LK1, node distance=1cm](LV1){$L_{v}$};
\node[draw, above of=LK1, node distance=1cm](LQ1){$L_{q}$};

\draw (X1b) -- (LQ1);
\draw (X1b) -- (LK1);
\draw (X1b) -- (LV1);

\node[right of=LQ1, node distance=1cm](Q1){$\mathbf{q}_0$};
\node[right of=LK1, node distance=1cm](K1){$\mathbf{k}_0$};
\node[right of=LV1, node distance=1cm](V1){$\mathbf{v}_0$};

\draw (LQ1) -- (Q1);
\draw (LK1) -- (K1);
\draw (LV1) -- (V1);

\coordinate[right of= X3, node distance=2cm](X3b){};

\draw (X3) -- (X3b);

\node[draw, right of=X3b, node distance=1cm](LK3){$L_{k}$};
\node[draw, below of=LK3, node distance=1cm](LV3){$L_{v}$};

\draw (X3b) -- (LK3);	
\draw (X3b) -- (LV3);

\node[right of=LK3, node distance=1cm](K3){$\mathbf{k}_{n}$};
\node[right of=LV3, node distance=1cm](V3){$\mathbf{v}_{n}$};

\draw (LK3) -- (K3);
\draw (LV3) -- (V3);

\coordinate[right of=Q1, node distance=0.3cm](TOP){};
\coordinate[right of=V3, node distance=0.3cm](BOT){};
\draw[decorate,decoration={brace}] (TOP) -- node[left=5pt]{} (BOT);

\node[right of=X2, text width=3cm, node distance=5.5cm](EQ){
};

\node[draw, right of=X2, node distance=8.3cm](EQ2){
	\footnotesize $\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)V$};

\coordinate[left of=EQ2, node distance=1.6cm, above=2.2cm](TOP2){};
\coordinate[left of=EQ2, node distance=1.5cm, below=0.0cm](MID2){};
\coordinate[left of=EQ2, node distance=1.6cm, below=2.2cm](BOT2){};
\draw[decorate,decoration={brace}] (TOP2) -- node[left=5pt]{} (BOT2);
\draw (MID2) -- (EQ2);

\node[right of=EQ2, node distance=2cm](OUT){output};
\draw (EQ2) -- (OUT);

\draw[draw=black, dashed] (1.75cm, -5.5cm) rectangle (9.5cm,1.5cm);
\end{tikzpicture}

\end{center}

}

%
%%----------------------------------------------------------------------------------------
%% Experiments
%%----------------------------------------------------------------------------------------
%
\headerbox{Experiments}{name=experiments,span=2,column=1,below=archi, above=bottom}{ 
\begin{center}
	\textbf{Agents}

	\begin{tabular}{lccc}
		\toprule
		Agent & \MLPL & \CNNG & \EgoAtt \\
		\midrule 
		Input sizes & [15, 7] & [11, 11, 7] & [1-15, 7] \\
		Layers sizes & [128, 128] &  \makecell[tc]{Convolutional layers: 3 \\ Kernel Size: 2 \\
			Stride: 2 \\ Head: [20]} & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads, $d_k=32$ \\ Decoder: [64, 64]} \\
		Number of parameters & 3.0e4 & 3.2e4 & 3.4e4 \\
		Variable input size & No & No &  {Yes}  \\
		Permutation invariant & No & {Yes} &  {Yes} \\
		\bottomrule
	\end{tabular}
\end{center}

\begin{center}
	\textbf{Performances}

	\includegraphics[width=0.32\linewidth]{../img/total_reward}
	\includegraphics[width=0.32\linewidth]{../img/length}
	\includegraphics[width=0.32\linewidth]{../img/velocity}
\end{center}

\begin{center}
\textbf{Interpretation}

\includegraphics[width=0.26\linewidth]{../img/head_specialization}
\includegraphics[width=0.28\linewidth]{../img/distances}
\includegraphics[width=0.21\linewidth]{../img/watch1}
\includegraphics[width=0.21\linewidth]{../img/watch2}

Head specialisation \qquad\qquad\qquad Distance \qquad\qquad\qquad\qquad\qquad Sensitivity to uncertainty \qquad\qquad\qquad\qquad

\includegraphics[width=.159\linewidth]{../img/episode1}
\includegraphics[width=.159\linewidth]{../img/episode2}
\includegraphics[width=.159\linewidth]{../img/episode3}
\includegraphics[width=.159\linewidth]{../img/episode4}
\includegraphics[width=.159\linewidth]{../img/episode5}
\includegraphics[width=.159\linewidth]{../img/episode6}

A full episode

\includegraphics[width=0.2\linewidth]{../img/priority1}
\includegraphics[width=0.2\linewidth]{../img/priority2}

Exploiting interaction patterns
\end{center}

}

%----------------------------------------------------------------------------------------
%	Acknowledgements
%----------------------------------------------------------------------------------------
\headerbox{Acknowledgements}{name=ack,column=0,span=1,below=setting}{
This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
}

%----------------------------------------------------------------------------------------
%	References
%----------------------------------------------------------------------------------------
\headerbox{References}{name=refs,column=0,span=1,below=ack, above=bottom}{
    {
    \AtNextBibliography{\footnotesize}
    \setlength{\bibitemsep}{3pt}
    \printbibliography[heading=none]
    }
}

\end{poster}

\end{document}