### 2019-11-27 paper/poster/img/architecture.tex
Update poster

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
% \usepackage[nonatbib]{neurips_2019_ml4ad}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
% \usepackage[preprint, nonatbib]{neurips_2019_ml4ad}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage[final, nonatbib]{neurips_2019_ml4ad}

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{neurips_2019_ml4ad}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

% Custom packages
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{array, makecell}
\usepackage{tikz}
\usepackage{xspace}
\usepackage{float}
\usetikzlibrary{arrows,automata}
\usepackage{pgfplots}
\usetikzlibrary{intersections}
\usepackage{subcaption}
\usepackage[flushleft]{threeparttable}
\usepackage{cases}
\usepackage{multirow}

\definecolor{mylinkcolor}{HTML}{e83f6f}
\definecolor{mycitecolor}{HTML}{2980B9}
\definecolor{myurlcolor}{HTML}{d63230}
% \definecolor{myalgocolor}{HTML}{32936f}
\definecolor{myalgocolor}{HTML}{000000}
\hypersetup{
	linkcolor  = mylinkcolor,
	citecolor  = mycitecolor,
	urlcolor   = myurlcolor,
	colorlinks = true,
}

\AtBeginDocument{\def\sectionautorefname{Section}}%

\input{mathdef}

\title{Social Attention for Autonomous Decision-Making in Dense Traffic}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

%Happy Horizontal People Transporter
%From the Sirius Cybernetics Corporation with a Genuine People Personality

\author{%
	Edouard Leurent\thanks{Equal contribution.} \\
	SequeL team, INRIA Lille -- Nord Europe\\
	Renault Group, France\\
	\texttt{edouard.leurent@inria.fr} \\
	\And
	Jean Mercat$^*$ \\
	Laboratoire des signaux et des syst\`emes, Centrale-Sup\'elec\\
	Renault Group, France\\
	\texttt{jean.mercat@renault.com} \\
}

\begin{document}

	\tikzset{
		state/.style={
			rectangle,
			draw=black, very thick,
			minimum height=2em,
			inner sep=2pt,
			text centered,
		},
		name plot/.style={every path/.style={name path global=#1}}
	}

		\begin{tikzpicture}
		\node(X1){ego};
		\node[below of=X1, node distance=0.7cm](X2){vehicle$_{1}$};
		\node[below of=X2, node distance=0.6cm](X3){$\vdots$};
		\node[below of=X3, node distance=0.7cm](X4){vehicle$_{N}$};

		\node[draw, right of=X1, node distance=1.8cm, rectangle](ENC1){Encoder};
		\node[draw, right of=X2, node distance=1.8cm, rectangle](ENC2){Encoder};
		\node[below of=ENC2, node distance=0.6cm](ENC3){$\vdots$};
		\node[draw, right of=X4, node distance=1.8cm, rectangle](ENC4){Encoder};

		\path (X1) edge (ENC1);
		\path (X2) edge (ENC2);
		\path (X4) edge (ENC4);

		\node[draw, rectangle, right of=ENC1, node distance=2.0cm, below=-0.4cm, fill=white](TRANS3){\rotatebox{90}{ Ego-attention }};
		\node[draw, rectangle, right of=ENC1, node distance=1.9cm, below=-0.3cm, fill=white](TRANS2){\rotatebox{90}{ Ego-attention }};
		\node[draw, rectangle, right of=ENC1, node distance=1.8cm, below=-0.2cm, fill=white](TRANS1){\rotatebox{90}{ Ego-attention }};

		\draw (ENC1.east) -| (TRANS1.west);
		\draw (ENC2.east) -| (TRANS1.west);
		\draw (ENC4.east) -| (TRANS1.west);

		\node[draw, right of=ENC1, node distance=3.6cm, rectangle](DEC1){Decoder};

		\draw (TRANS1.east) |- (DEC1.west);

		\node[right of=DEC1, node distance=1.5cm](Y1){$Q$};

		\draw (DEC1.east) -- (Y1.west);
		\end{tikzpicture}

		\clearpage
		\begin{tikzpicture}[scale=1, every node/.style={scale=1}]
		\node(X1){ego encoding};
		\node[below of=X1, node distance=2cm](X2){$\vdots$};
		\node[below of=X2, node distance=2cm](X3){vehicle$_{N}$ encoding};

		\coordinate[right of= X1, node distance=2cm](X1b){};

		\draw (X1) -- (X1b);

		\node[draw, right of=X1b, node distance=1cm](LK1){$L_{k}$};
		\node[draw, below of=LK1, node distance=1cm](LV1){$L_{v}$};
		\node[draw, above of=LK1, node distance=1cm](LQ1){$L_{q}$};

		\draw (X1b) -- (LQ1);
		\draw (X1b) -- (LK1);
		\draw (X1b) -- (LV1);

		\node[right of=LQ1, node distance=1cm](Q1){$\mathbf{q}_0$};
		\node[right of=LK1, node distance=1cm](K1){$\mathbf{k}_0$};
		\node[right of=LV1, node distance=1cm](V1){$\mathbf{v}_0$};

		\draw (LQ1) -- (Q1);
		\draw (LK1) -- (K1);
		\draw (LV1) -- (V1);

		\coordinate[right of= X3, node distance=2cm](X3b){};

		\draw (X3) -- (X3b);

		\node[draw, right of=X3b, node distance=1cm](LK3){$L_{k}$};
		\node[draw, below of=LK3, node distance=1cm](LV3){$L_{v}$};

		\draw (X3b) -- (LK3);	
		\draw (X3b) -- (LV3);

		\node[right of=LK3, node distance=1cm](K3){$\mathbf{k}_{n}$};
		\node[right of=LV3, node distance=1cm](V3){$\mathbf{v}_{n}$};

		\draw (LK3) -- (K3);
		\draw (LV3) -- (V3);

		\coordinate[right of=Q1, node distance=0.3cm](TOP){};
		\coordinate[right of=V3, node distance=0.3cm](BOT){};
		\draw[decorate,decoration={brace}] (TOP) -- node[left=5pt]{} (BOT);

		\node[right of=X2, text width=3cm, node distance=5.5cm](EQ){
			\footnotesize \[Q = \left( \begin{matrix}
			\mathbf{q}_0
			\end{matrix} \right)\]
			\\
			\footnotesize \[K = \left( \begin{matrix}
			\mathbf{k}_0 \\
			\vdots \\
			\mathbf{k}_{n}
			\end{matrix} \right)\]
			\\
			\footnotesize \[ V = \left( \begin{matrix}
			\mathbf{v}_0 \\
			\vdots \\
			\mathbf{v}_{n}
			\end{matrix}\right) \]
		};

		\node[draw, right of=X2, node distance=8.1cm](EQ2){
			\footnotesize $\underbrace{\sigma\left(\frac{QK^T}{\sqrt{d_k}}\right)}_{\text{attention matrix}}V$};

		\coordinate[left of=EQ2, node distance=1.6cm, above=2.2cm](TOP2){};
		\coordinate[left of=EQ2, node distance=1.5cm, below=0.0cm](MID2){};
		\coordinate[left of=EQ2, node distance=1.6cm, below=2.2cm](BOT2){};
		\draw[decorate,decoration={brace}] (TOP2) -- node[left=5pt]{} (BOT2);
		\draw (MID2) -- (EQ2);

		\node[right of=EQ2, node distance=2cm](OUT){output};
		\draw (EQ2) -- (OUT);

		\draw[draw=black, dashed] (1.75cm, -5.5cm) rectangle (9.5cm,1.5cm);
		\end{tikzpicture}

\end{document}