### 2019-11-27 paper/poster/poster.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% baposter Landscape Poster
% LaTeX Template
% Version 1.0 (15/5/13)
%
% Created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a0paper,portrait,fontscale=0.3, margin=2cm]{baposter}

\usepackage[utf8]{inputenc}
\usepackage[font=normalsize,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\captionsetup{justification=centering}

\usepackage{booktabs} % Horizontal rules in tables
\usepackage{lmodern}
\usepackage{relsize} % Used for making text smaller in some places

% Other packages
\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}}
\def\imagetop#1{\vtop{\null\hbox{#1}}}
\usepackage{adjustbox}
\usepackage{subfig}
\usepackage{floatrow} % sidecapfloat
\usepackage{setspace}
\usepackage{xspace}
\usepackage{noReferences}
\usepackage{tikz}
\usepackage[customcolors]{hf-tikz}
\usepackage{epsf}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{float}
\usepackage{cases}
\usepackage{verbatim}
\usepackage{array}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{multirow}
\usepackage{listings}
%\usepackage[authoryear]{natbib}
\usepackage{authblk}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{eqparbox}
\usepackage{xcolor}
\usepackage[backend=biber, citestyle=authoryear, maxcitenames=2, maxbibnames=2]{biblatex}
\usepackage{lipsum}
\usepackage{svg}
\usepackage{cleveref}
\usepackage{tikz}
\usepackage{xspace}
\usetikzlibrary{arrows,automata}
\usepackage{array, makecell}
\usepackage{pgfplots}
\usetikzlibrary{intersections}
\usepackage{pifont}

\bibliography{../budgeted_rl.bib}

\input{style}
\input{../mathdef}

\begin{document}

%\background{ % Set the background to an image (background.pdf)
%\begin{tikzpicture}[remember picture,overlay]
%\draw (current page.north west)+(-2em,2em) node[anchor=north west]
%{\includegraphics[height=1.1\textheight]{background.png}};
%\end{tikzpicture}
%}
\begin{poster}{
grid=false,
borderColor=bordercol, % Border color of content boxes
headerColorOne=headercol1, % Background color for the header in the content boxes (left side)
headerColorTwo=headercol2, % Background color for the header in the content boxes (right side)
headerFontColor=headerfontcol, % Text color for the header text in the content boxes
boxColorOne=boxcolor, % Background color for the content in the content boxes
headershape=roundedright, % Specify the rounded corner in the content box headers
%headerfont=\Large\sf\bf, % Font modifiers for the text in the content box headers
headerfont=\Large\bf\textsc, %Sans Serif
textborder=rectangle,
background=none,
headerborder=open, % Change to closed for a line under the content box headers
boxshade=plain,
textfont={\setlength{\parindent}{0.0em}\sffamily},
headerheight={0.05\textheight},
eyecatcher=true
%columns=5
}
%
%----------------------------------------------------------------------------------------
%	Title and authors
%----------------------------------------------------------------------------------------
%
{
\includegraphics[width=10em]{./img/companies}
}
{
Social Attention for Autonomous Decision-Making in Dense Traffic
}
{
Edouard Leurent, Jean Mercat
\vspace{-4\baselineskip}
}
{
\includegraphics[width=12em]{./img/inria_sc}
}

\setlength{\colheight}{0.92\textheight}

%----------------------------------------------------------------------------------------
%	Motivation
%----------------------------------------------------------------------------------------

\headerbox{\textsc{Motivation}}{name=motivation,span=1,column=0,row=0}{

\textbf{Behavioural Planning}

\begin{itemize}
	\item \hlg{High-level} decisions: lane, stop or go
	\item In \hlr{interaction} with other drivers
\end{itemize}

\textbf{Reinforcement Learning}

\begin{equation*}
Q^*(s, a) \eqdef \max_\pi \expectedvalue\left[\sum_{t=0}^\infty \gamma^t R(s_t, a_t)\condbar \substack{
s_0=s\\a_0=a
}
\right]
\end{equation*}

Bellman Optimality Equation:
\begin{equation*}
Q^*(s, a) = \expectedvalueover{s'\sim P(s'|s, a)} \max_{a'\in A} \left[R(s, a) + \gamma Q^*(s', a')\right]
\end{equation*}
}

%----------------------------------------------------------------------------------------
%	Setting
%----------------------------------------------------------------------------------------
\headerbox{\textsc{State Representation}}{name=setting,span=1,column=0,row=0,below=motivation}{

\textbf{Representing social interactions}

A joint state $s$ of $N+1$ observed vehicles
\begin{align*}
s &= \left( s_i \right)_{i \in [0, N]}\\
s_i &= \begin{bmatrix}
x_i & y_i & v^x_i & v^y_i & \cos\psi_i & \sin \psi_i
\end{bmatrix}^T
\label{eq:coordinates}
\end{align*}

\begin{center}
	\includegraphics[width=0.35\textwidth]{../img/coordinates}
	\captionof*{figure}{List of features representation}
\end{center}

\textbf{Issues related to function approximation}

\begin{enumerate}
	\item \hlg{Variable} size
    \begin{itemize}
		\item[\incarrow] usual models accept \hlr{fixed-size} inputs
	\end{itemize}

	\item \hlr{Sensitivity} to the ordering
	\begin{itemize}
		\item we want the policy to be \hlg{permutation-invariant}:
	\end{itemize}
\begin{align*}
\forall\tau \in \mathfrak{S}_N, \qquad &\pi(\cdot|(s_0, s_1,\dotsc,s_N)) = \\ &\pi(\cdot|(s_0, s_{\tau(1)},\dotsc,s_{\tau(N)}))
\end{align*}
\end{enumerate}

\textbf{A common solution}
\begin{center}
	\includegraphics[width=0.35\textwidth]{../img/map}
	\captionof*{figure}{Occupancy grid representation}
\end{center}

\begin{itemize}[nolistsep]
	\item[\hlg{\checkmark}] Fixed-size
	\item[\hlg{\checkmark}] Does not depend on an ordering
	\item[\hlr{\xmark}] Suffers from an \hlr{accuracy / size tradeoff}
\end{itemize}

}

%----------------------------------------------------------------------------------------
%	Model Architecture
%----------------------------------------------------------------------------------------
\headerbox{Model Architecture}{name=archi,column=1,span=2}{

\begin{center}
	\begin{minipage}{0.4\textwidth}
		\includegraphics[trim={3.8cm 22.5cm 10cm 2.5cm}, clip, page=1, width=\linewidth]{img/architecture}
		\captionof*{figure}{Model architecture}
		\begin{itemize}
			\item[\hlg{\checkmark}] Inputs can have a {variable size}
			\item[\hlg{\checkmark}] Based on a dot product
			\begin{itemize}
				\item[$\incarrow$] {permutation-invariant}
			\end{itemize}
			\item[\hlg{\checkmark}] Compact size with no accuracy loss
		\end{itemize}
	\end{minipage}
	\begin{minipage}{0.59\textwidth}
		\includegraphics[trim={3.8cm 18.3cm 5.8cm 2.5cm}, clip, page=2, width=\linewidth]{img/architecture}
		\captionof*{figure}{Ego-attention block}
	\end{minipage}

\end{center}

}

%
%%----------------------------------------------------------------------------------------
%% Experiments
%%----------------------------------------------------------------------------------------
%
\headerbox{Experiments}{name=experiments,span=2,column=1,below=archi, above=bottom}{ 
\begin{center}
	\captionof*{table}{\textbf{Agents}}
	\begin{tabular}{lccc}
		\toprule
		Agent & \MLPL & \CNNG & \EgoAtt \\
		\midrule 
		Input sizes & [15, 7] & [32, 32, 7] & [~$\boldsymbol{\cdot}$~, 7] \\
		Layers sizes & [128, 128] &  \makecell[tc]{Convolutional layers: 3 \\ Kernel Size: 2 \\
			Stride: 2 \\ Head: [20]} & \makecell[tl]{Encoder: [64, 64] \\Attention: 2 heads\\\phantom{Attention: }$d_k=32$ \\ Decoder: [64, 64]} \\
		Number of parameters & 3.0e4 & 3.2e4 & 3.4e4 \\
		Variable input size & \hlr{No} & \hlr{No} &  \hlg{Yes}  \\
		Permutation invariant & \hlr{No} & \hlg{Yes} &  \hlg{Yes} \\
		\bottomrule
	\end{tabular}
\end{center}

\begin{center}
	\captionof*{figure}{\textbf{Performances}}
	\includegraphics[width=0.32\linewidth]{../img/total_reward}
	\includegraphics[width=0.32\linewidth]{../img/velocity}
	\includegraphics[width=0.32\linewidth]{../img/length}
\end{center}

\begin{center}
\textbf{Attention visualisation}

\begin{minipage}[b]{0.26\textwidth}
\includegraphics[width=\linewidth]{../img/head_specialization}
\captionof*{figure}{Head specialisation}
\end{minipage}
\begin{minipage}[b]{0.28\textwidth}
	\includegraphics[width=\linewidth]{../img/distances}
\captionof*{figure}{Distance}
\end{minipage}
\begin{minipage}[b]{0.42\textwidth}
	\includegraphics[width=0.49\linewidth]{../img/watch1}
	\includegraphics[width=0.49\linewidth]{../img/watch2}
	\captionof*{figure}{Sensitivity to uncertainty}
\end{minipage}

\begin{center}
\includegraphics[width=.159\linewidth]{../img/episode1}
\includegraphics[width=.159\linewidth]{../img/episode2}
\includegraphics[width=.159\linewidth]{../img/episode3}
\includegraphics[width=.159\linewidth]{../img/episode4}
\includegraphics[width=.159\linewidth]{../img/episode5}
\includegraphics[width=.159\linewidth]{../img/episode6}
\captionof*{figure}{A full episode}
\end{center}

\includegraphics[width=0.18\linewidth]{../img/priority1}
\includegraphics[width=0.18\linewidth]{../img/priority2}

Exploiting interaction patterns
\end{center}

}

%----------------------------------------------------------------------------------------
%	Acknowledgements
%----------------------------------------------------------------------------------------
\headerbox{Acknowledgements}{name=ack,column=0,span=1,below=setting, above=bottom}{
This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
}

%----------------------------------------------------------------------------------------
%	References
%----------------------------------------------------------------------------------------
%\headerbox{References}{name=refs,column=0,span=1,below=ack, above=bottom}{
%    {
%    \AtNextBibliography{\footnotesize}
%    \setlength{\bibitemsep}{3pt}
%    \printbibliography[heading=none]
%    }
%}

\end{poster}

\end{document}